---
title: "Problem2"
author: "Srikar"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

### Clear all Outputs and Terminal of R Studio

```{r}
## This for safety and to avoid variables conflict
rm(list = ls())
graphics.off()
cat("\014")
```

### Installing Packages Required

```{r message=FALSE, warning=FALSE}

options (repos = "https://cran.rstudio.com")
packages <- c ("ggplot2","readxl","plotly","stargazer", "regclass","leaps", "car")
installed_packages <- packages %in% rownames(installed.packages())

# Install packages, if not yet installed
if (any(installed_packages == FALSE)) {
install.packages ( packages [!installed_packages])
}

invisible (lapply(packages,library,character.only = TRUE))   # Package Loading
```

### Loading Data

```{r}
getwd()
obs <- read_excel("Problem2set2.xlsx")
```

```{r}
### Number of Columns
ncol(obs)
### Number of rows
n_obs=nrow(obs)
```

```{r}
head(obs)
Y_obs <- obs[, 1]$'Yvariable'
factors <- obs[, 2:6]

FactorA <- factors[, 1]$'FactorA'
FactorB <- factors[, 2]$'FactorB'
FactorC <- factors[, 3]$'FactorC'
FactorD <- factors[, 4]$'FactorD'
FactorE <- factors[, 5]$'FactorE'
```

### Shapiro-Wilk Normality Test

```{r}
tests <- list(
  "Factor A" = shapiro.test(FactorA),
  "Factor B" = shapiro.test(FactorB),
  "Factor C" = shapiro.test(FactorC),
  "Factor D" = shapiro.test(FactorD),
  "Factor E" = shapiro.test(FactorE)
)

tests
```

### Kolmogorov-Smirnov

```{r}
## Test for Normal Distribution
ks.test(FactorA,pnorm,mean=mean(FactorA),sd=sd(FactorA))
ks.test(FactorB,pnorm,mean=mean(FactorB),sd=sd(FactorB))
ks.test(FactorC,pnorm,mean=mean(FactorC),sd=sd(FactorC))
ks.test(FactorD,pnorm,mean=mean(FactorD),sd=sd(FactorD))
ks.test(FactorE,pnorm,mean=mean(FactorE),sd=sd(FactorE))

```

### Histogram Plots

Bins are choosed based on Freedman–Diaconis (FD) rule, for robustness towards to outliners

```{r}
fig <- plot_ly(alpha = 0.6)
# fig <- fig %>% add_histogram(x = FactorA,marker = list(line = list(color = "red", width = 1)))
fig <- fig %>% add_histogram(x = FactorA, name = 'Factor A')
fig <- fig %>% add_histogram(x = FactorB, name = 'Factor B')
fig <- fig %>% add_histogram(x = FactorC, name = 'Factor C')
fig <- fig %>% add_histogram(x = FactorD, name = 'Factor D')
fig <- fig %>% add_histogram(x = FactorE, name = 'Factor E')
fig <- fig %>% layout(barmode = "overlay")

fig
```

```{r}
hist(FactorA,
     probability = TRUE,
     breaks = "FD",
     main = "Factor A",
     xlab = "Value")

lines(density(FactorA), lwd = 2)

x <- seq(min(FactorA), max(FactorA), length.out = 200)
lines(x,
      dnorm(x, mean(FactorA), sd(FactorA)),
      col = "red",
      lwd = 2)

legend("topright",
       legend = c("Kernel Density Estimate", "Normal Fit"),
       lwd = c(2, 2),
       col = c("black", "red"),
       bty = "n")
```

```{r}
hist(FactorB,
     probability = TRUE,
     breaks = "FD",
     main = "Factor B",
     xlab = "Value")

lines(density(FactorB), lwd = 2)

x <- seq(min(FactorB), max(FactorB), length.out = 200)
lines(x,
      dnorm(x, mean(FactorB), sd(FactorB)),
      col = "orange",
      lwd = 2)

legend("topright",
       legend = c("Kernel Density Estimate", "Normal Fit"),
       lwd = c(2, 2),
       col = c("black", "orange"),
       bty = "n")
```

```{r}
hist(FactorC,
     probability = TRUE,
     breaks = "FD",
     main = "Factor C",
     xlab = "Value")

lines(density(FactorC), lwd = 2)

x <- seq(min(FactorC), max(FactorC), length.out = 200)
lines(x,
      dnorm(x, mean(FactorC), sd(FactorC)),
      col = "forestgreen",
      lwd = 2)

legend("topright",
       legend = c("Kernel Density Estimate", "Normal Fit"),
       lwd = c(2, 2),
       col = c("black", "forestgreen"),
       bty = "n")
```

```{r}
hist(FactorD,
     probability = TRUE,
     breaks = "FD",
     main = "Factor D",
     xlab = "Value")

lines(density(FactorD), lwd = 2)

x <- seq(min(FactorD), max(FactorD), length.out = 200)
lines(x,
      dnorm(x, mean(FactorD), sd(FactorD)),
      col = "blue",
      lwd = 2)

legend("topright",
       legend = c("Kernel Density Estimate", "Normal Fit"),
       lwd = c(2, 2),
       col = c("black", "blue"),
       bty = "n")
```

```{r}
hist(FactorE,
     probability = TRUE,
     breaks = "FD",
     main = "Factor E",
     xlab = "Value")

lines(density(FactorE), lwd = 2)

x <- seq(min(FactorE), max(FactorE), length.out = 200)
lines(x,
      dnorm(x, mean(FactorE), sd(FactorE)),
      col = "purple",
      lwd = 2)

legend("topright",
       legend = c("Kernel Density Estimate", "Normal Fit"),
       lwd = c(2, 2),
       col = c("black", "purple"),
       bty = "n")
```

### Quantile-Quantile Plots

```{r}
qq_plot <- ggplot() +
  stat_qq(aes(sample = FactorA, color = "FactorA")) +
  stat_qq(aes(sample = FactorB, color = "FactorB")) +
  stat_qq(aes(sample = FactorC, color = "FactorC")) +
  stat_qq(aes(sample = FactorD, color = "FactorD")) +
  stat_qq(aes(sample = FactorE, color = "FactorE")) +
  stat_qq_line(aes(sample = FactorA, color = "FactorA")) +
  stat_qq_line(aes(sample = FactorB, color = "FactorB")) +
  stat_qq_line(aes(sample = FactorC, color = "FactorC")) +
  stat_qq_line(aes(sample = FactorD, color = "FactorD")) +
  stat_qq_line(aes(sample = FactorE, color = "FactorE")) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    color = "Factor",
    title = "Q-Q Plot of Factors"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qq_plot
```

```{r}
qqplotA <- ggplot(data.frame(FactorA), aes(sample = FactorA)) +
  stat_qq() +
  stat_qq_line(color = "red", linewidth = 1) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Factor A"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qqplotA
```

```{r}
qqplotB <- ggplot(data.frame(FactorB), aes(sample = FactorB)) +
  stat_qq() +
  stat_qq_line(color = "orange", linewidth = 1) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Factor B"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qqplotB
```

```{r}
qqplotC <- ggplot(data.frame(FactorC), aes(sample = FactorC)) +
  stat_qq() +
  stat_qq_line(color = "forestgreen", linewidth = 1) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Factor C"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qqplotC
```

```{r}
qqplotD <- ggplot(data.frame(FactorD), aes(sample = FactorD)) +
  stat_qq() +
  stat_qq_line(color = "blue", linewidth = 1) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Factor D"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qqplotD
```

```{r}
qqplotE <- ggplot(data.frame(FactorE), aes(sample = FactorE)) +
  stat_qq() +
  stat_qq_line(color = "purple", linewidth = 1) +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles",
    title = "Factor E"
  ) +
  scale_x_continuous(n.breaks = 10) +
  scale_y_continuous(n.breaks = 10) +
  theme_light()

qqplotE
```

### Simple Linear Regression Models

```{r}
# Simple Linear Regression For Factor A
regA <- lm(Y_obs~FactorA)
summary(regA)
```

```{r}
# Simple Linear Regression For Factor B
regB <- lm(Y_obs~FactorB)
summary(regB)
```

```{r}
# Simple Linear Regression For Factor C
regC <- lm(Y_obs~FactorC)
summary(regC)
```

```{r}
# Simple Linear Regression For Factor D
regD <- lm(Y_obs~FactorD)
summary(regD)
```

```{r}
# Simple Linear Regression For Factor E
regE <- lm(Y_obs~FactorE)
summary(regE)
```

#### Summary to LaTeX table

Using stargazer package, we are getting the simple regression model summary as LaTeX table in the output. **Un-comment** and Change the parameter regA for different factors.

```{r}
## Regression summary to LaTeX table
stargazer(
  regE,
  type = "latex",
  title = "Regression Results on Factor E",
  label = "tab:regE_summ",
  single.row = TRUE,
  dep.var.labels = "Yvariable",
  covariate.labels = c("Factor E")
)
```

### Multi-Linear Regression Model

```{r}
multireg <- lm(Y_obs~FactorA+FactorB+FactorC+FactorD+FactorE)
summary(multireg)
anova(multireg)
```

```{r}
# stargazer(multireg, type = "latex",
#           title = "Multi Regression Results",
#           label = "tab:multireg_summ",
#           single.row = TRUE)
```

### Multi-Collinearity Test on Factors

```{r}
vif(multireg)
```

### Bayes Information Criterion

```{r}
BIC(multireg)
```

Choosing the best model using, step function, a stepwise model selection for regression models. criteria="AIC" for Akaike Information Criterion. direction= "both" meaning allowing both addition and removal of factors based on the goodness of model

```{r}
best_model <- step(multireg, direction="both", k=log(n_obs))
summary(best_model)
BIC(best_model)
```

```{r}
stargazer(best_model, type = "latex",
          title = "Multi Regression Results of Best Model",
          label = "tab:bestmodel_summ",
          single.row = TRUE)
```

### Best Model based on Adjusted R\^2- Squared

```{r}
regfit <- regsubsets(
  Y_obs ~ FactorA + FactorB + FactorC + FactorD + FactorE, data=obs,
  nvmax = n_obs
)
summary(regfit)
```

```{r}
adjr2_plot <- plot(regfit, scale="adjr2", col=gray(seq(0, 0.9, length = 10)))
title(main="Adjusted R²", font.main=2)
BIC_plot <- plot(regfit, scale="bic")
title(main="BIC", font.main=2)
```

### Prediction using the Multi-Linear Regression Model

```{r}
# Creating a data frame of new factors and calling the column "new_factors".
###This naming of the column is important whenn using predict afterwards. 
new_factors <- data.frame(
  FactorA = 34,
  FactorB = 76,
  FactorC = -25,
  FactorD = -55,
  FactorE = 42
)

# Predicting the new value of the Yvaribale
new_Y_obs <-predict(best_model,newdata=new_factors)
new_Y_obs
```
